Технология использования открытых данных Федеральной налоговой службы России для экономико-географического анализа

### 2024-04-23 legal-companies-geography/Paper.Rmd
Edit paragraphs about Figure 1 and Fugure 2, modify Figure 2, make lexical changes, #1

This paper provides an exploratory regional analysis of legal business in Russia for 2016–21. The analysis is based mainly on the Federal Tax Service (FTS) open data on small and medium-sized businesses and on the number of employees of organisations, as well as on portions of official statistical information. The paper proposes a reproducible technique of working with FTS open data for regional analysis in service geography that was partially implemented in a specially designed Python command-line tool. The analysis includes questions about the count of legal firms and their employees by region, about the share of organisations in regional legal business, about the role of legal business in region's economy, and about the yearly change of the number of legal companies in regions. In addition, it briefly addresses an issue of probable connection between the location of multi-regional courts and the development of legal business. It is found out that legal business in Russia is highly concentrated in Moscow and Saint Petersburg that jointly account for 1/3 of legal firms and workers and are also at the top by the role of legal business in the region. However, there are other regions with relatively high development of legal business, that are located mainly on the European South of Russia, in the Volga-Urals and in Siberia. There is a group of regions with domination of organisations in legal business and the group of regions where individual entrepreneurs prevail over companies. District courts are usually located in the regions with high level of legal business development, but the mechanism of this co-location requires further investigation. In general, the number of legal firms in regions grown from 2016 to 2021, with the common inverted U-shape pattern probably explained by the coronavirus (covid-19) pandemic that negetively impacted the small and medium business and reverted the initial growth trend.

@@

The research adheres to the general framework of regional analysis and employs the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. The analysis is comprised of four research problems, or questions, which are enumerated below with the description of the relevant techniques used to find the answers. The term “legal firm” denotes both companies (organizations) and individuals (individual entrepreneurs). The term “legal workers” refers to the employees of these firms.

Problem 1 is the distribution of legal firms across Russia's regions. To discover this distribution, the median number of legal firms in each region is calculated for the period from 2016 to 2021. This result is then visualized on a map.

Problem 2 is the structure of legal business in the regions with respect to the kind of firms (either individuals or companies). To address this issue, the median share of legal companies in each region over the 2016–2021 period is calculated and then mapped.

Problem 3 is the relative importance of regional legal business which is divided into two sub-problems. The first is the importance of a region's legal services in the Russian legal market measured by the share of legal workers in a region in the number of country's legal workers. The second is the importance of these services within a given region estimated as a share of legal workers in a region in the total number of workers in that region. Both metrics are calculated for each year from 2018 to 2021, averaged by a median function, and then plotted on a tile-grid map. Additionally, regions with district-level courts are marked on the map to analyze the relationship between the location of such courts and the development of legal business.

Problem 4 is the temporal change in the legal business sector across regions. To reach conclusions, legal firms are calculated by region and year for the 2016–2021 period. The time series of counts are visualized as line plots to identify temporal patterns. The regions are divided into the three groups based on the overall growth, decline, or stability of legal business in the observed period, and the group is also marked on the map.

@@

**Average yearly count of legal firms by region** is displayed on a tile-grid colorpleth map in Fig. 1. Several conclusions can be drawn from this visualisation. First, Moscow is the only city where the number of legal firms is above 10 thousand, and this is the highest value in Russia. Next, a set of regions also has relatively high count of legal companies and individual entrepreneurs. In spatial terms, most of these regions form several groups. One of these groups is located at the south of the European part of Russia and consists of Krasnodar and Stavropol territories with Rostov, Volgograd and Voronezh regions. Tatarstan and Bashkir republics with Samara, Sverdlovsk, Chelyabinsk regions and Perm territory make up another group that may be called Ural-Volga. Some Siberian regions, namely Tyumen, Omsk, Novosibirsk, Kemerovo, with Krasnoyarsk territory, form the third group that may be named Siberian by its location. Also, Saint Petersburg city, Nizhniy Novgorod region, and Primorie territory act as single-region groups. The majority of other regions has relatively low count of legal firms, and in some of them (several republics of North Caucas and Altai and a portion of regions from the North) there are less than a hundred of such companies. The co-loocation of regions and groups with different count of legal firms suggests that there may be a sort of central-peripheries relations between them. For example, Moscow city and region, Saint Petersburg city, and Nizhniy Novgorod region are surrounded by dozens of regions with relatively small number of legal firms, and thus may act as a centers of legal business for them. The same may be true for the pairs of South and Caucasian, as well as for Siberia and Far East. It may be hypothesized that economic development of regions is positively associated with the development of legal business. Also, the population of a region may also matter.

@@

```{r fig-1-counts-map, fig.cap="Fig. 1. Median yearly count of legal firms (both organisations and individual entrepreneurs) by regions of Russia for 2016–21. Grey is no data"}
counts_map
```

@@

Fig. 2 displays the *median shares of organisations among all legal firms in regions*. Legal firms may be either organisations or individual entrepreneurs, so for each region, the rest part of the firms consists of individuals. The share of companies among all legal firms varies from less about 25% to almost 70%. However, a substantial part of regions has from 40% to 60% of organisations, and thus many regions may be categorized either as with a small domination of companies or as with small domination of individuals on the legal market. Several southern regions (Krasnodar and Stavropol territories, Rostov region, Kalmyk republic), as well as several northern ones (Arkhangelsk region, Komi republic, Nenetz and Yamal territories), and some central regions (Kaluga and Moscow) have substantially lower shares of companies compared to the surrounding regions, thus forming the areas of predominantly individual legal business. The overall count of legal firms seems to have to connection with the share of organisation among them.

@@

```{r fig-2-shares-map, fig.cap="Fig. 2. Median yearly shares of companies among legal firms by regions of Russia for 2016–21. Grey is no data"}
shares_map
```

>>>>

This paper provides an exploratory regional analysis of legal business in Russia for 2016–21. The analysis is based mainly on the Federal Tax Service (FTS) open data on small and medium-sized businesses and on the number of employees of organisations, as well as on portions of official statistical information. The paper proposes a reproducible technique of working with FTS open data for regional analysis in service geography that was partially implemented in a specially designed Python command-line tool. The analysis includes questions about the count of law firms and their employees by region, about the share of organisations in regional legal business, about the role of legal business in region's economy, and about the yearly change of the number of legal companies in regions. In addition, it briefly addresses an issue of probable connection between the location of multi-regional courts and the development of legal business. It is found out that legal business in Russia is highly concentrated in Moscow and Saint Petersburg that jointly account for 1/3 of legal firms and workers and are also at the top by the role of legal business in the region. However, there are other regions with relatively high development of legal business, that are located mainly on the European South of Russia, in the Volga-Urals and in Siberia. There is a group of regions with domination of organisations in legal business and the group of regions where individual entrepreneurs prevail over companies. District courts are usually located in the regions with high level of legal business development, but the mechanism of this co-location requires further investigation. In general, the number of legal firms in regions grown from 2016 to 2021, with the common inverted U-shape pattern probably explained by the coronavirus (covid-19) pandemic that negetively impacted the small and medium business and reverted the initial growth trend.

@@

The research adheres to the general framework of regional analysis and employs the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. The analysis is comprised of four research problems, or questions, which are enumerated below with the description of the relevant techniques used to find the answers. The term “law firm” denotes both companies (organizations) and individuals (individual entrepreneurs). The term “lawyers” in this paper refers to the employees of these firms.

Problem 1 is the distribution of law firms across Russia's regions. To discover this distribution, the median number of law firms in each region is calculated for the period from 2016 to 2021. This result is then visualized on a map.

Problem 2 is the structure of legal business in the regions with respect to the kind of firms (either individuals or companies). To address this issue, the median share of law firms in each region over the 2016–2021 period is calculated and then mapped.

Problem 3 is the relative importance of regional legal business which is divided into two sub-problems. The first is the importance of a region's legal services in the Russian legal market measured by the share of lawyers in a region in the number of country's lawyers. The second is the importance of these services within a given region estimated as a share of lawyers in a region in the total number of employees in that region. Both metrics are calculated for each year from 2018 to 2021, averaged by a median function, and then plotted on a tile-grid map. Additionally, regions with district-level courts are marked on the map to analyze the relationship between the location of such courts and the development of legal business.

Problem 4 is the temporal change in the legal business sector across regions. To reach conclusions, law firms are calculated by region and year for the 2016–2021 period. The time series of counts are visualized as line plots to identify temporal patterns. The regions are divided into the three groups based on the overall growth, decline, or stability of legal business in the observed period, and the group is also marked on the map.

Technically, the analysis is carried out using the R programming language (including additional packages). A tile grid map of Russia by I. Dianov[^10] is used with a modified implementation by Yu. Tukachev[^11]. The source code of this paper is published in the author's repository on GitHub[^12].

[^10]: Плиточная карта России. Web resource: https://dianov.org/all/plitochnaya-karta-rossii/ (accessed 23.04.2024)
[^11]: Плиточная карта РФ. Web resource: https://gist.github.com/tukachev/c3262242a20b3050e0459b4e3afc3a51 (accessed 23.04.2024)
[^12]: PavelSyomin / ru-smb-companies-papers. Web resource: https://github.com/PavelSyomin/ru-smb-companies-papers/tree/main/legal-companies-geography (accessed 23.04.2024)

@@

Figure 1 depicts the average annual number of law firms by region. Several conclusions can be drawn from this visualization. First, Moscow is the only city where the number of law firms exceeds 10 thousand, and this is the highest value in Russia. Second, some regions also have a relatively high number of law firms. Geographically, these regions form several groups. One of these groups is located in the south of the European part of Russia and consists of Krasnodar and Stavropol Territories with Rostov, Volgograd and Voronezh regions. Tatarstan and Bashkir Republics with Samara, Sverdlovsk, and Chelyabinsk Regions, as well as Perm Territory form the Ural-Volga group. Tyumen, Omsk, Novosibirsk, Kemerovo Regions with Krasnoyarsk Territory belong to the third group which can be designated as Siberian due to its location. The city of St. Petersburg, Nizhniy Novgorod Region, and Primorye Territory are separate regions with relatively high number of law firms. Third, the majority of other regions have a relatively small number of law firms, with several republics of the North Caucasus and Altai and a portion of regions of the Russian North incorporating less than a hundred of such companies. It can be hypothesized that the number of law firms is positively associated with the economic development of a region. Furthermore, the population of a region may also be a factor to consider.

The  relative location of regions and groups with different numbers of law firms suggests the possibility of central-peripheries relations between them. For instance, Moscow City and Region, St. Petersburg City, and Nizhny Novgorod Region may serve as hubs for their surrounding regions of the Central and North-Western Russia. A similar phenomenon may be observed in other parts of the country: the Siberian group provides legal services for the Far East, while the regions of the European South serve the republics of the North Caucasus.

@@

```{r fig-1-counts-map, fig.cap="Fig. 1. The median annual count of law firms by regions of Russia for the period from 2016 to 2021"}
counts_map
```

@@

Figure 2 illustrates the median annual share of organizations among all law firms in regions. The value varies considerably, ranging from approximately 30% to almost 70%. However, a substantial proportion of regions exhibit a share of companies ranging from 45% to 55%, which is almost equal to the share of individuals. All other regions may be divided into two broad groups: either having a dominance of individuals or organizations in the legal market. The former category encompasses the European North and South, as well as certain regions in the Fart East, Central Russia, and Western Siberia. The latter category comprises Moscow and St. Petersburg cities, along with the majority of the Volga or Siberian regions. The comparison of Figure 1 with Figure 2 leads to the conclusion that there is no clear relationship between the number of law firms and the share of companies among them.

@@

```{r fig-2-shares-map, fig.cap="Fig. 2. The median annual share of companies among legal firms by regions of Russia for the period from 2016 to 2021"}
shares_map
```
